High-accuracy neutron diffusion calculations based on integral transport theory

### 2019-12-12 docs/EPJP2019/Gross-EPJP-2019.tex
Further minor text fixes towards final version.

In 2004, Ronen~\cite{Ronen-2004} suggested to derive corrected diffusion coefficients using Fick's law and more accurate estimations of the neutron currents by means of integral transport operators, whereas the neutron flux is resolved by diffusion theory. Denoting the integral transport and diffusion currents by $\jtr\rE$ and $\bm{J}_D\rE$, respectively, the Ronen idea is based on changing the diffusion coefficient such that
\begin{equation}\label{eq:Fick}
\bm{J}_D\rE = -D\rE\bnabla\phi\rE = \jtr\rE. 
\end{equation}

@@

Instead of computing new diffusion coefficients at the interfaces by Fick's law, $\jd(x,E) = -D(E) \partial_x \phi(x,E)$, new corrective currents $\jd(x_s,E) = \jtrr(x_s,E) - \jd(x_s,E)$ are obtained at cell interfaces $x_s$. Here, $\jd(x,E)$ is called the ``diffusion current'' and is obtained using Fick's law, with the original values of the diffusion coefficients and with the derivative approximated by finite differences. In one-dimensional 
geometry and using the notation in \fig{fig:mesh1D}, the diffusion current and the corrective current are evaluated according to
\begin{subequations}\label{eq:JD-dJ}
\begin{align}
\jd(x_{i+1/2},E) \cong& -D(x_{i+1/2},E) 
\frac{\phi(x_{i+1},E) - \phi(x_i,E)}
{(\Delta x_{i+1} + \Delta x_i)/2}, \label{eq:JD} \\
\delta J(x_{i+1/2},E) =& -\delta D(x_{i+1/2},E) 
\frac{\phi(x_{i+1},E) + \phi(x_i,E)}
{(\Delta x_{i+1} + \Delta x_i)/2}, \label{eq:dJ}
\end{align}
\end{subequations}
where integer and rational subscripts indicate node-averaged and interface quantities, respectively. The discretized form of the current $\delj$ must involve the neighboring flux as well, but its representation is changed into a drift-advection term to get rid of possible undefined division by zeros in case of flat flux~\cite{Smith-1983,Tomatis-2011}. The so-called ``transport current'' $\jtrr(x_s,E)$ is evaluated using the integral expression derived in~\sxn{sec:int-flux-J} as described in~\sxn{sec:calc-int-curr}.

@@

Using the definitions in~\eqsand{eq:JD-dJ}{eq:Ds} with the notation shown in~\fig{fig:mesh1D}, the discretized form of~\eq{eq:diff1} is
\begin{equation}\label{eq:FD_diff_full}
\begin{split}
- 2D_{i+1/2}^g\frac{\phi_{i+1,g}-\phi_{i,g}}{\dx_{i+1} + \dx_i}
- 2\delta D_{i+1/2}^g\frac{\phi_{i+1,g}+\phi_{i,g}}{(\dx_{i+1}+\dx_i)}
&+ 2D_{i-1/2}^g\frac{\phi_{i,g}-\phi_{i-1,g}}{\dx_{i} + \dx_{i-1}}  
+ 2\delta D_{i-1/2}^g\frac{\phi_{i,g}+\phi_{i-1,g}}{(\dx_i+\dx_{i-1})} \\
&+ \sigma_{i}^g \phi_{i,g} 
= \sum\limits_{g'=1}^G \sigma_{s,i}^{g\leftarrow g'}\phi_{i,g'}
+ \frac{\chi_g}{\keff}\sum\limits_{g'=1}^G
 \nu\sigma_{f,i}^{g'}\phi_{i,g'}.
\end{split}
\end{equation}
Rearranging terms, one gets
\begin{equation}\label{eq:FD_diff_org}
\begin{split}
\left(
\frac{-D_{i-1/2}^g + \delta D_{i-1/2}^g }{\dx_{i-1}+\dx_{i}}
\right) \phi_{i-1,g}
&+ \left[
\frac{D_{i+1/2}^g - \delta D_{i+1/2}^g}{\dx_{i+1}+\dx_i}
+\frac{D_{i-1/2}^g + \delta D_{i-1/2}^g}{\dx_{i}+\dx_{i-1}}
+\frac{\sigma_{i}^g}{2}
\right] \phi_{i,g} \\
&+ \left(
\frac{-D_{i+1/2}^g -\delta D_{i+1/2}^g }{\dx_{i+1}+\dx_{i}}
\right) \phi_{i+1,g} = \frac{1}{2}q_{i,g},
\end{split}
\end{equation}
where $q_{i,g}$ is the RHS of~\eq{eq:FD_diff_full}.

@@

The Ronen algorithm is described as follows (see Fig.~\ref{fig:RM-algo}):
\begin{enumerate}
	\item\label{it:init} Initialize - the algorithm receives the geometry and the cross sections (XS) of the problem as input. At this stage, the correction factors $\delta D^{(0)}$ are set to zero and the initial pure-diffusion operator $\mathcal{A}_0$ is constructed. Initial guesses for the flux $\phi^{(0)}$ and multiplication factor $\keff^{(0)}$ are set.
	\item\label{it:Js} Calculate currents - using the diffusion flux $\phi^{(k)}$, both the diffusion current $\jd^{(k)}$ (Eq.~\ref{eq:JD-dJ}) and the transport current $\jtrr^{(k)}$ (Eq.~\ref{eq:tot-J}) are calculated.
	\item\label{it:dD} Calculate correction factors - the correction factors $\delta D^{(k+1)}$ (Eq.~\ref{eq:JD-dJ}) is obtained using $\delj^{(k)}$ and Eq.~\eqref{eq:JD-dJ}.
	\item\label{it:dA} Update diffusion operator - reconstruct the diffusion operator $\mathcal{A}^{(k+1)} = \mathcal{A}_0+\delta \mathcal{A}^{(k+1)}$ (Eq.~\ref{eq:FD_diff_org}).
	\item\label{it:DF-solve} Diffusion solver - a one-dimensional multigroup diffusion solver is executed until convergence. The results are an updated estimation of the diffusion flux $\phi^{(k+1)}$ and the multiplication factor $\keff^{(k+1)}$.
	\item\label{it:cvg} Check convergence - check if the flux $\phi^{(k)}$ and the multiplication factor $\keff^{(k)}$ converged. If yes, stop, otherwise, step~\ref{it:Js}.
\end{enumerate}

\begin{figure}[htbp!] 
	\centering
\begin{tikzpicture}[font=\scriptsize]
% Place nodes
	\node [cloud] 					 						(start) {START};
	\node [block, below of=start, xshift=0cm,yshift=1.5cm]  (input) {Input\\parameters};
	\node [block, below of=input, xshift=0cm,yshift=1.5cm]  (geoXS) {Geometry, XS};
	\node [block, below of=geoXS, xshift=0cm,yshift=1.5cm]  (J)     {Calculate $\jtrr^{(k)},\jd^{(k)},\delj^{(k)}$};
	\node [block, below of=J,     xshift=0cm,yshift=1.5cm]  (dDA)   {Obtain $\delta D^{(k+1)}$, $\delta \mathcal{A}^{(k+1)}$};
	\node [block, below of=dDA,  xshift=0cm,yshift=1.5cm]  	(Anew)  {Construct\\$\mathcal{A}_0+\delta \mathcal{A}^{(k+1)}$};
	\node [block, below of=Anew, xshift=0cm,yshift=1.5cm]  	(Diff)  {Diffusion\\solver};
	\node [decision, below of=Diff,xshift=0cm,yshift=-1cm]	(cvg)   {$\phi^{(k+1)}$, $k^{(k+1)}$ \\ converged?};
	\node [cloud, below of=cvg,xshift=0cm,yshift=1.5cm] 	(stp)  {STOP};

% Draw edges
	\path [line] (start) -- (input);
	\path [line] (input) -- (geoXS);
	\path [line] (geoXS) -- node[anchor= south west,yshift=-0.2cm]{$\phi^{(0)}$, $k^{(0)}$, $\mathcal{A}_0$}(J);
	\path [line] (J) 	 -- (dDA);
	\path [line] (dDA) 	 -- (Anew);
	\path [line] (Anew)  -- node[near end,anchor= south west,xshift=0cm]{$\mathcal{A}^{(k+1)}$}(Diff);
	\path [line] (Diff)  -- node[anchor= south west,yshift=-0.2cm]{$\phi^{(k+1)},k^{(k+1)}$}(cvg);
	\path [line] (cvg)   -- node[anchor= south west]{YES}(stp);
	\path [line] (cvg)   -| node[near start,yshift=0.2cm]{NO}(-3,-6.0) -- node[near end]{}(J);

@@

Two one-dimensional test cases are considered, homogeneous and heterogeneous. The reference solutions are calculated using a Python3 discrete ordinates $S_N$ code with $N=16$~\cite{Lewis-1984}. The diffusion solutions and the Ronen iterations are produced using original codes still developed in Python3. All tolerances on relative residuals are set to 1E-6 in the iterative solving schemes and no (over/under)-relaxation is done. The initial diffusion coefficient is set to $D=1/(3\sigma_{tr}) = 1/(3\sigma)$ since we only consider here isotropic scattering. Anisotropic scattering cases are deferred to future studies. Void boundary conditions are used at the edges of the slab, which are implemented for the diffusion solver in the form of a group-dependent extrapolated distance, i.e., $d_g = 2.13 D_g$, see~\sxn{sec:BC}.

@@

A homogeneous slab of width $a=21.5$ cm is considered, with the two-group macroscopic cross sections shown in Table~\ref{tab:xs}. 
%
\begin{table}[htbp]
	\centering
	\caption{Two-group macroscopic cross sections [cm\tsup{-1}] of the homogeneous case~\cite{Tomatis-2011}.}
	\label{tab:xs}
	\begin{tabular}{cccccc}
		%\midrule
		Group $g$ &  $\sigma_{g}$ & \multicolumn{2}{c}{$\sigma_{s,0,g\leftarrow g'}$} & $\chi_g$ & $\nu\sigma_{f,g}$ \\ 
		\midrule
		1 & 5.3115$\times$10\tsup{-1} & 5.04664$\times$10\tsup{-1} & 2.03884$\times$10\tsup{-3} & 1 & 7.15848$\times$10\tsup{-3} \\
		%\midrule
		2 & 1.30058$\times$10\tsup{+0}& 1.62955$\times$10\tsup{-2} & 1.19134$\times$10\tsup{+0}	& 0 & 1.41284$\times$10\tsup{-1} \\
		%\midrule
	\end{tabular}
\end{table}

Table~\ref{tab:dx_drho} shows the reactivity differences as a function of mesh refinement using standard diffusion (with extrapolated length) and the Ronen method. The reactivity difference is given by $(\Delta\rho = 1/\kref - 1/k_\texttt{D/RM})\times 10^5$, where $\kd$ and $\krm$ are the multiplication factors of standard diffusion and of the Ronen method, respectively. The reference $\kref$ value from S\tsub{16} is 0.744417 ($I = 400$).

\begin{table}[htbp]
	\centering
	\caption{Reactivity differences as a function of mesh refinement using the Ronen method.}
	\label{tab:dx_drho}
	\begin{tabular}{lcccccr}
		%\midrule
		$dx$[cm] & $I$ &  $\kref$ & $\kd$ & $\Delta\rho_\texttt{D}$ [pcm] & $\krm$ & $\Delta\rho_\texttt{RM}$ [pcm]\\ 
		\midrule
		0.43 & ~50     & 0.744307 & 0.741417 & -524 & 0.740552  & -681\\
		%\midrule
		0.215 & 100    & 0.744391 & 0.741355 & -550 & 0.743447 & -171\\
		%\midrule
		0.1075 & 200   & 0.744412 & 0.741339 & -557 & 0.744212 & -36\\
		%\midrule
		0.07167 & 300 & 0.744416 & 0.741336 & -558 & 0.744356 & -11\\
		%\midrule
		0.05375 & 400  & 0.744417 & 0.741335 & -558 & 0.744407 & -2\\
		%\midrule
	\end{tabular}
\end{table}

A comparison of the fast and thermal fluxes, as calculated by the reference $S_N$ code, the RM code, and the standard multigroup diffusion without RM correction ($D_0$), is shown in Fig.~\ref{fig:slab-fluxes} for the half slab. The deviations (in [\%]) of the RM-corrected flux and of the standard diffusion flux from the reference $S_N$ flux are also shown in Fig.~\ref{fig:slab-fluxes-dev}. The corresponding Ronen method correction factors $\delta D$ are shown in Fig.~\ref{fig:slab-RM-corr-factor}.

The deviation from the reference solution is decreased by the Ronen iterations from approximately 20\% near the boundary and 1\% at the slab center to 2\% and 0\%, respectively.

@@

\begin{figure}[htbp]
	\centering
	\includegraphics[width=0.7\linewidth]{Sn_Diff_Rm_flux_Tomatis2011.pdf}
	\caption{Comparison of the fluxes as calculated by the reference $S_N$ code, the RM code, and a standard multigroup diffusion without RM correction ($D_0$). Results shown here after 250 Ronen iterations.}
	\label{fig:slab-fluxes}		
\end{figure}

@@

The RM correction terms ($\delta D$) along the slab are shown in Fig.~\ref{fig:slab-RM-corr-factor}. It is clear that the diffusion correction assumes zero values at the slab center and exhibit non-trivial behavior near the boundary. The convergence of the flux (max deviation) and of the criticality eigenvalue are shown in Fig.~\ref{fig:conv}. While the eigenvalue converges fast within a few iterations, the flux exhibit slow convergence in general with higher differences against the reference $S_{16}$ solution near the boundaries. The spatial flux convergence between two successive Ronen iterations is shown in Fig.~\ref{fig:conv2} for the left half slab, according to $\Delta \phi = (\phi^{k-1}-\phi^k)/\phi^{k-1}$ [\%]. The spatial flux convergence of the Ronen iterations with respect to the reference solution for the left half slab is shown in Fig.~\ref{fig:conv3} instead.

@@

% convergence
\begin{figure}[hbtp]
	\centering
	\includegraphics[width=0.48\linewidth]{flux_deviation_half_it_1_sn.pdf}
	\includegraphics[width=0.48\linewidth]{flux_deviation_half_it_2_sn.pdf}
	\caption{The spatial flux convergence of the Ronen iterations with respect to the reference $S_N$ solution for the left half slab.}
	\label{fig:conv3}
\end{figure}

Notice the change of sign in the flux convergence $\Delta \phi$ in Fig.~\ref{fig:conv2} between the center (negative) and the boundary (positive) of the slab. The initial diffusion solution is smoother than the transport one. Hence, it can be seen in Fig.~\ref{fig:slab-fluxes} that the diffusion solution underestimates the peak flux at the center of the slab and overestimates the flux near the boundary. Ronen iterations ``push'' the diffusion solution towards the transport one by increasing the flux level at the center of the slab and decreasing the flux level at the boundary. Hence, $\phi^{(k-1)}$ is smaller at the center and higher at the boundary compared to $\phi^{(k)}$. This effect caused by Ronen iterations produces less smooth flux shape with steeper gradients comparing to the diffusion solution.
%
% ---------------------------------------------------
\subsection{Heterogeneous case}
\label{subsec:heterog}

@@

	\node[above] at (-2,1.5) {Assembly 1};

@@

	\node[above] at (1,1.5) {Assembly 2};

@@

	\node[above] at (-2,-1) {Assembly 3};

@@

	\draw [<->] (-3,-2.7) -- node[below]{$15.24[cm]$}	(-1,-2.7);

@@

	\node[above] at (1,-1) {Assembly 4};

@@

	\draw [<->] (4,-2.8) -- node[below]{$106.68[cm]$}	(11,-2.8);

@@

\begin{table}[htbp!]	
	\centering
	\caption{Two-group macroscopic cross sections [cm\tsup{-1}] used for the heterogeneous case~\cite{Rahnema-1997}. Colors are in accordance with Fig.~\ref{fig:hetro-core}.} 
	\label{tab:xs2}
	\begin{tabular}{lcccccccr}
		Material&$\sigma_1$&$\sigma_2$&$\nu\sigma_{f,1}$&$\nu\sigma_{f,2}$&$\sigma_{s,1\leftarrow 1}$&$\sigma_{s,2\leftarrow 1}$&$\sigma_{s,2\leftarrow 2}$&$x[cm]$\\
		\midrule
		Water  (Blue)	&0.1890	 &1.4633  &0.0000	 &0.0000	 &0.1507   &0.0380		&1.4536		 &1.158\\
		Fuel 1 (Red)	&0.2263  &1.0119  &0.0067	 &0.1241	 &0.2006   &0.0161		&0.9355		 &3.231\\
		Fuel 2 (Yellow)	&0.2252  &0.9915  &0.0078	 &0.1542 	 &0.1995   &0.0156		&0.9014		 &3.231\\
		Fuel 3 (Green)	&0.2173  &1.0606  &0.0056	 &0.0187	 &0.1902   &0.0136		&0.5733 	 &3.231\\
	\end{tabular}
\end{table}

Table~\ref{tab:dx_drho2} shows the reactivity differences between standard diffusion (with extrapolated length) and the Ronen method (at increasing number of iterations) with respect to the reference solution for each of the three core configurations. The standard diffusion solution perform rather well for core 1 with deviation of 182 pcm, but fails for cores 2 and 3 with deviation of 773 and 3335 pcm, respectively. This is explained by the fact that cores 2 and 3 are more heterogeneous with respect to core 1. The Ronen method performs better than standard diffusion for all three cores, even after merely 10 iterations. Additional Ronen iterations further improve the estimated multiplication factor but less with milder changes.

@@

Numerically, Tomatis and Dall'Osso 2011 chose to deal with this potential singularity by using the well-known drift terms, which serve as a \emph{numerical feature}. The (more accurate) surface current $\jtrr$, calculated by integral expression, is written as the sum of the surface current calculated by diffusion, $\jd$, plus a correction $\delj$. Hence, when the current vanishes, the correction $\delj$ vanishes as well. Moreover, the correction is written as in Eq.~\eqref{eq:JD-dJ}, avoiding the potential divergence resulting from division by zero.

>>>>

In 2004, Ronen~\cite{Ronen-2004} suggested to derive corrected diffusion coefficients using Fick's law and more accurate estimations of the neutron currents by means of integral transport operators, whereas the neutron flux is resolved by diffusion theory. Denoting the currents obtained by integral transport as ``integral currents'' $\jtr\rE$ and the currents obtained by Fick's law as ``diffusion currents'' $\bm{J}_D\rE$, the Ronen idea is based on changing the diffusion coefficient such that
\begin{equation}\label{eq:Fick}
\bm{J}_D\rE = -D\rE\bnabla\phi\rE = \jtr\rE. 
\end{equation}

@@

Instead of computing new diffusion coefficients at the interfaces by Fick's law, $\jd(x,E) = -D(E) \partial_x \phi(x,E)$, new corrective currents $\delta J(x_s,E) = \jtrr(x_s,E) - \jd(x_s,E)$ are obtained at cell interfaces $x_s$. Here, $\jd(x,E)$ is called the ``diffusion current'' and is obtained using Fick's law, with the original values of the diffusion coefficients and with the derivative approximated by finite differences. In one-dimensional geometry and using the notation in \fig{fig:mesh1D}, the diffusion current is evaluated according to
\begin{equation}\label{eq:JD}
\jd(x_{i+1/2},E) \cong -D(x_{i+1/2},E) 
\frac{\phi(x_{i+1},E) - \phi(x_i,E)}
{(\Delta x_{i+1} + \Delta x_i)/2},
\end{equation}
and the corrective current is defined according to
\begin{equation}\label{eq:dJ}
\delta J(x_{i+1/2},E) = -\delta D(x_{i+1/2},E) 
\frac{\phi(x_{i+1},E) + \phi(x_i,E)}
{(\Delta x_{i+1} + \Delta x_i)/2},
\end{equation}
where integer and rational subscripts indicate node-averaged and interface quantities, respectively. It is Eq.~\eqref{eq:dJ} from which the correction factors $\delta D$ are obtained. The discretized form of the current $\delj$ must involve the neighboring flux as well, but its representation is changed into a drift-advection term to get rid of possible undefined division by zeros in case of flat flux~\cite{Smith-1983,Tomatis-2011}. The so-called ``transport current'' $\jtrr(x_s,E)$ is evaluated using the integral expression derived in~\sxn{sec:int-flux-J} as described in~\sxn{sec:calc-int-curr}.

@@

Using the definitions in Eqs.~\eqsthru{eq:JD}{eq:Ds} with the notation shown in~\fig{fig:mesh1D}, the discretized form of~\eq{eq:diff1} is
\begin{equation}\label{eq:FD_diff_full}
\begin{split}
- 2D_{i+1/2}^g\frac{\phi_{i+1,g}-\phi_{i,g}}{\dx_{i+1} + \dx_i}
- 2\delta D_{i+1/2}^g\frac{\phi_{i+1,g}+\phi_{i,g}}{(\dx_{i+1}+\dx_i)}
&+ 2D_{i-1/2}^g\frac{\phi_{i,g}-\phi_{i-1,g}}{\dx_{i} + \dx_{i-1}}  
+ 2\delta D_{i-1/2}^g\frac{\phi_{i,g}+\phi_{i-1,g}}{(\dx_i+\dx_{i-1})} \\
&+ \sigma_{i}^g \phi_{i,g} 
= \sum\limits_{g'=1}^G \sigma_{s,i}^{g\leftarrow g'}\phi_{i,g'}
+ \frac{\chi_g}{\keff}\sum\limits_{g'=1}^G
 \nu\sigma_{f,i}^{g'}\phi_{i,g'}.
\end{split}
\end{equation}
Rearranging terms, one gets
\begin{equation}\label{eq:FD_diff_org}
\begin{split}
\left(
\frac{-D_{i-1/2}^g + \delta D_{i-1/2}^g }{\dx_{i-1}+\dx_{i}}
\right) \phi_{i-1,g}
&+ \left[
\frac{D_{i+1/2}^g - \delta D_{i+1/2}^g}{\dx_{i+1}+\dx_i}
+\frac{D_{i-1/2}^g + \delta D_{i-1/2}^g}{\dx_{i}+\dx_{i-1}}
+\frac{\sigma_{i}^g}{2}
\right] \phi_{i,g} \\
&+ \left(
\frac{-D_{i+1/2}^g -\delta D_{i+1/2}^g }{\dx_{i+1}+\dx_{i}}
\right) \phi_{i+1,g} = \frac{1}{2}q_{i,g},
\end{split}
\end{equation}
where $q_{i,g}$ is the RHS of~\eq{eq:FD_diff_full}.

@@

The Ronen algorithm is described as follows (see Fig.~\ref{fig:RM-algo}):
\begin{enumerate}
	\item\label{it:init} Initialize - the algorithm receives the geometry and the cross sections (XS) of the problem as input. At this stage, the correction factors $\delta D^{(0)}$ are set to zero and the initial pure-diffusion operator $\mathcal{A}_0$ is constructed. Initial guesses for the flux $\phi^{(0)}$ and multiplication factor $\keff^{(0)}$ are set.
	\item\label{it:Js} Calculate currents - using the flux $\phi^{(k)}$ obtained from the diffusion solver, both the diffusion current $\jd^{(k)}$ (Eq.~\ref{eq:JD}) and the transport current $\jtrr^{(k)}$ (Eq.~\ref{eq:tot-J}) are calculated.
	\item\label{it:dD} Calculate correction factors - the correction factors $\delta D^{(k+1)}$ are obtained using $\delj^{(k)}$ and Eq.~\eqref{eq:dJ}.
	\item\label{it:dA} Update the diffusion operator - reconstruct the diffusion operator $\mathcal{A}^{(k+1)} = \mathcal{A}_0+\delta \mathcal{A}^{(k+1)}$ (Eq.~\ref{eq:FD_diff_org}).
	\item\label{it:DF-solve} Diffusion solver - a one-dimensional multigroup diffusion solver is executed until convergence. The results are an updated estimation of the flux $\phi^{(k+1)}$ and the multiplication factor $\keff^{(k+1)}$.
	\item\label{it:cvg} Check convergence - check if the flux $\phi^{(k)}$ and the multiplication factor $\keff^{(k)}$ converged. If yes, stop, otherwise, step~\ref{it:Js}.
\end{enumerate}

\begin{figure}[htbp!] 
	\centering
\begin{tikzpicture}[font=\scriptsize]
% Place nodes
	\node [cloud] 					 						(start) {START};
	\node [block, below of=start, xshift=0cm,yshift=2.2cm]  (geoXS) {Geometry, XS};
	\node [block, below of=geoXS, xshift=0cm,yshift=1.8cm]  (J)     {Calculate $\jtrr^{(k)},\jd^{(k)},\delj^{(k)}$};
	\node [block, below of=J,     xshift=0cm,yshift=2.0cm]  (dDA)   {Obtain $\delta D^{(k+1)}$ $\delta \mathcal{A}^{(k+1)}$};
	\node [block, below of=dDA,  xshift=0cm,yshift=2.0cm]  	(Anew)  {Construct\\$\mathcal{A}_0+\delta \mathcal{A}^{(k+1)}$};
	\node [block, below of=Anew, xshift=0cm,yshift=1.8cm]  	(Diff)  {Diffusion\\solver};
	\node [decision, below of=Diff,xshift=0cm,yshift=-.5cm]	(cvg)   {$\phi^{(k+1)}$ $k_\textrm{eff}^{(k+1)}$ \\ converged?};
	\node [cloud, below of=cvg,xshift=0cm,yshift=1.8cm] 	(stp)  {STOP};

% Draw edges
	\path [line] (start) -- (geoXS);
	\path [line] (geoXS) -- node[anchor= south west,yshift=-0.2cm]{$\phi^{(0)}$, $k_\textrm{eff}^{(0)}$, $\mathcal{A}_0$}(J);
	\path [line] (J) 	 -- (dDA);
	\path [line] (dDA) 	 -- (Anew);
	\path [line] (Anew)  -- node[near end,anchor= south west,xshift=0cm]{$\mathcal{A}^{(k+1)}$}(Diff);
	\path [line] (Diff)  -- node[anchor= south west,yshift=-0.2cm]{$\phi^{(k+1)},k_\textrm{eff}^{(k+1)}$}(cvg);
	\path [line] (cvg)   -- node[anchor= south west,yshift=-0.1cm]{YES}(stp);
	\path [line] (cvg)   -| node[near start,yshift=0.2cm]{NO}(-3,-3.0) -- node[near end]{}(J);

@@

Two one-dimensional test cases are considered, homogeneous and heterogeneous. The reference solutions are calculated using a Python3 discrete ordinates $S_N$ code with $N=16$~\cite{Lewis-1984}. The diffusion solutions and the Ronen iterations are produced using original codes in Python3. All tolerances on relative residuals are set to 1E-6 in the iterative solving schemes and no (over/under)-relaxation is done. The initial diffusion coefficient is set to $D=1/(3\sigma_{tr}) = 1/(3\sigma)$ since only isotropic scattering is considered in this case. Anisotropic scattering cases are deferred to future studies. Void boundary conditions are used at the outer edges of the core, which are implemented for the diffusion solver in the form of a group-dependent extrapolated distance, i.e., $d_g = 2.13 D_g$, see~\sxn{sec:BC}.

@@

A homogeneous slab of width $a=21.5$ cm is considered, with the two-group macroscopic cross sections shown in Table~\ref{tab:xs}. 
%
\begin{table}[htbp]
	\centering
	\caption{Two-group macroscopic cross sections [cm\tsup{-1}] of the homogeneous case~\cite{Tomatis-2011}.}
	\label{tab:xs}
	\begin{tabular}{cccccc}
		%\midrule
		Group $g$ &  $\sigma_{g}$ & \multicolumn{2}{c}{$\sigma_{s,0,g\leftarrow g'}$} & $\chi_g$ & $\nu\sigma_{f,g}$ \\ 
		\midrule
		1 & 5.3115$\times$10\tsup{-1} & 5.04664$\times$10\tsup{-1} & 2.03884$\times$10\tsup{-3} & 1 & 7.15848$\times$10\tsup{-3} \\
		%\midrule
		2 & 1.30058$\times$10\tsup{0}& 1.62955$\times$10\tsup{-2} & 1.19134$\times$10\tsup{0}	& 0 & 1.41284$\times$10\tsup{-1} \\
		%\midrule
	\end{tabular}
\end{table}

Table~\ref{tab:dx_drho} shows the reactivity differences as a function of mesh refinement using standard diffusion (with extrapolated length) and the Ronen method. The reactivity difference is given by $(\Delta\rho = 1/\kref - 1/k_\textrm{D/RM})\times 10^5$ pcm, where $\kd$ and $\krm$ are the multiplication factors of standard diffusion and of the Ronen method, respectively. The reference $\kref$ value from S\tsub{16} is 0.744417 ($I = 400$).

\begin{table}[htbp]
	\centering
	\caption{Reactivity differences as a function of mesh refinement using the Ronen method.}
	\label{tab:dx_drho}
	\begin{tabular}{lcccccr}
		%\midrule
		$dx$[cm] & $I$ &  $\kref$ & $\kd$ & $\Delta\rho_\textrm{D}$ [pcm] & $\krm$ & $\Delta\rho_\textrm{RM}$ [pcm]\\ 
		\midrule
		0.43 & ~50     & 0.744307 & 0.741417 & -524 & 0.740552  & -681\\
		%\midrule
		0.215 & 100    & 0.744391 & 0.741355 & -550 & 0.743447 & -171\\
		%\midrule
		0.1075 & 200   & 0.744412 & 0.741339 & -557 & 0.744212 & -36\\
		%\midrule
		0.07167 & 300 & 0.744416 & 0.741336 & -558 & 0.744356 & -11\\
		%\midrule
		0.05375 & 400  & 0.744417 & 0.741335 & -558 & 0.744407 & -2\\
		%\midrule
	\end{tabular}
\end{table}

A comparison of the fast and thermal fluxes, as calculated by the reference $S_N$ code, the RM code, and the standard multigroup diffusion without RM correction (denoted by $D_0$), is shown in Fig.~\ref{fig:slab-fluxes} for the half slab. The deviations (in [\%]) of the RM-corrected flux and of the standard diffusion flux from the reference $S_N$ flux are also shown in Fig.~\ref{fig:slab-fluxes-dev}. The corresponding Ronen method correction factors $\delta D$ are shown in Fig.~\ref{fig:slab-RM-corr-factor}.

The deviation from the reference solution is decreased by the Ronen iterations from approximately 20\% near the boundary and 1\% at the slab center to 2\% and 0.01\%, respectively.

@@

\begin{figure}[htbp]
	\centering
	\includegraphics[width=0.7\linewidth]{Sn_Diff_Rm_flux_Tomatis2011.pdf}
	\caption{Comparison of the fluxes as calculated by the reference $S_N$ code, the RM code, and a standard multigroup diffusion without RM correction ($D_0$). The insets show a zoom-in of the fluxes near the boundary. Results shown here after 250 Ronen iterations.}
	\label{fig:slab-fluxes}		
\end{figure}

@@

The RM correction terms ($\delta D$) along the slab are shown in Fig.~\ref{fig:slab-RM-corr-factor}. It is clear that the diffusion correction assumes near-zero values at the slab center and exhibits non-trivial behavior near the boundary. The convergence of the flux (max deviation) and of the criticality eigenvalue are shown in Fig.~\ref{fig:conv}. While the eigenvalue converges within a few iterations, the flux exhibit relatively slow convergence with higher differences against the reference $S_{16}$ solution near the boundaries. The spatial flux convergence between two successive Ronen iterations is shown in Fig.~\ref{fig:conv2} for the left half slab, according to $\Delta \phi = (\phi^{k-1}-\phi^k)/\phi^{k-1}$ [\%]. The spatial flux convergence of the Ronen iterations with respect to the reference solution for the left half slab is shown in Fig.~\ref{fig:conv3} instead.

@@

% convergence
\begin{figure}[hbtp]
	\centering
	\includegraphics[width=0.48\linewidth]{flux_deviation_half_it_1_sn.pdf}
	\includegraphics[width=0.48\linewidth]{flux_deviation_half_it_2_sn.pdf}
	\caption{The spatial flux convergence of the Ronen iterations with respect to the reference $S_{16}$ solution for the left half slab.}
	\label{fig:conv3}
\end{figure}

Notice the change of sign in the flux convergence $\Delta \phi$ in Fig.~\ref{fig:conv2} between the center (negative) and the boundary (positive) of the slab. The initial diffusion solution is smoother than the transport one. Hence, it can be seen in Fig.~\ref{fig:slab-fluxes} that the diffusion solution underestimates the peak flux at the center of the slab and overestimates the flux near the boundary. The Ronen iterations ``push'' the diffusion solution towards the transport one by increasing the flux level at the center of the slab and decreasing the flux level at the boundary. Hence, $\phi^{(k-1)}$ is smaller at the center and higher at the boundary compared to $\phi^{(k)}$. This effect caused by Ronen iterations produces less smooth flux shape with steeper gradients comparing to the diffusion solution.
%
% ---------------------------------------------------
\subsection{Heterogeneous case}
\label{subsec:heterog}

@@

	\node[above] at (-2,1.5) {Assembly A1};

@@

	\node[above] at (1,1.5) {Assembly A2};

@@

	\node[above] at (-2,-1) {Assembly A3};

@@

	\draw [<->] (-3,-2.7) -- node[below]{15.24 [cm]}	(-1,-2.7);

@@

	\node[above] at (1,-1) {Assembly A4};

@@

	\draw [<->] (4,-2.8) -- node[below]{106.68 [cm]}	(11,-2.8);

@@

\begin{table}[htbp]	
	\centering
	\caption{Two-group macroscopic cross sections [cm\tsup{-1}] used for the heterogeneous case~\cite{Rahnema-1997}. Colors are in accordance with Fig.~\ref{fig:hetro-core}.} 
	\label{tab:xs2}
	\begin{tabular}{lcccccccr}
		Material&$\sigma_1$&$\sigma_2$&$\nu\sigma_{f,1}$&$\nu\sigma_{f,2}$&$\sigma_{s,1\leftarrow 1}$&$\sigma_{s,2\leftarrow 1}$&$\sigma_{s,2\leftarrow 2}$&$x$ [cm]\\
		\midrule
		Water  (Blue)	&0.1890	 &1.4633  &0.0000	 &0.0000	 &0.1507   &0.0380		&1.4536		 &1.158\\
		Fuel 1 (Red)	&0.2263  &1.0119  &0.0067	 &0.1241	 &0.2006   &0.0161		&0.9355		 &3.231\\
		Fuel 2 (Yellow)	&0.2252  &0.9915  &0.0078	 &0.1542 	 &0.1995   &0.0156		&0.9014		 &3.231\\
		Fuel 3 (Green)	&0.2173  &1.0606  &0.0056	 &0.0187	 &0.1902   &0.0136		&0.5733 	 &3.231\\
	\end{tabular}
\end{table}

%\newpage
Table~\ref{tab:dx_drho2} shows the reactivity differences between standard diffusion (with extrapolated length) and the Ronen method (at increasing number of iterations) with respect to the reference solution for each of the three core configurations. The standard diffusion solution perform rather well for core 1 with deviation of 182 pcm, but fails for cores 2 and 3 with deviation of 773 and 3335 pcm, respectively. This is explained by the fact that cores 2 and 3 are more heterogeneous with respect to core 1. The Ronen method performs better than standard diffusion for all three cores, even after merely 10 iterations. Additional Ronen iterations further improve the estimated multiplication factor until convergence.

@@

Numerically, Tomatis and Dall'Osso 2011 chose to deal with this potential singularity by using the well-known drift terms, which serve as a \emph{numerical feature}. The (more accurate) surface current $\jtrr$, calculated by integral expression, is written as the sum of the surface current calculated by diffusion, $\jd$, plus a correction $\delj$. Hence, when the current vanishes, the correction $\delj$ vanishes as well. Moreover, the correction is written as in Eq.~\eqref{eq:dJ}, avoiding the potential divergence resulting from division by zero.
